Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-01-22 Manuscript/manuscript.Rmd
manuscript edits

Replicate Hi-C data for the GM12878, K562, ... cell lines were downloaded from GEO GSE63525 (Experiments HIC001 to HIC018, HIC069 to HIC 074, ... respectively; produced using the MbolI restriction enzyme) [@rao20143d]. The autosomal genomic coordinates (GRCh37/hg19 human genome assembly) of TAD boundaries were obtained from data at 10 kb, 25 kb, 50 kb, and 100 kb resolution using the Arrowhead algorithm [@durand2016juicebox]. The start and end coordinates of all TADs were concatenated, sorted, and unique TAD boundary points were obtained. Genomic annotation data consisting of various chromatin segmentation states, histone modifications, and transcription factor binding sites (GRCh37/hg19 human genome assembly) were obtained from the UCSC Genome Browser Database [@Tyner:2017aa] (Supplementary Table S1).

@@

An ensemble of models was established for the purposes of systematically comparing the performances of different parameters in the framework. Each system in the framework is assumed to be composed of a model constructed from a particular data resolution, with a particular predictor type for the feature space (Figure 2).

@@

@@

>>>>

Replicate Hi-C data for the GM12878, K562, ... cell lines were downloaded from GEO GSE63525 (Experiments HIC001 to HIC018, HIC069 to HIC 074, ... respectively; produced using the MbolI restriction enzyme) [@rao20143d]. The autosomal genomic coordinates (GRCh37/hg19 human genome assembly) of TAD boundaries were obtained from data at 10 kb, 25 kb, 50 kb, and 100 kb resolution using the Arrowhead algorithm [@durand2016juicebox]. The start and end coordinates of all TADs were concatenated, sorted, and unique TAD boundary points were obtained.

@@

An ensemble of models was established for the purposes of systematically comparing the performances of different parameters in the framework (Figure 2). Each system in the framework is assumed to be composed of a model constructed from a particular data resolution, with a particular predictor type for the feature space.

@@

## DISTANCE TYPE PREDICTORS OUTPERFORM OTHER PREDICTOR-TYPES

We first evaluated which predictor types were best able to discriminate between genomic bins that either flanked or did not flank TAD boundaries. To do so we performed a series of random forest classification algorithms on the full features space (all combined predictor types), as well as on type-specific feature spaces (overlap counts, overlap predictors, and log2 distances). It was found that models with distance type predictors had higher AUPRC values compared to other predictor types (Figure 4). This result was observed across both data resolution and cell line, with the largest AUPRC obtained from the 10 kb resolutions for all [gm12878, k562, ???update as more cell lines included] cell lines (0.431 and 0.456 respectively). There was a gradual decline in AUPRC values as resolution increased with models built on 100 kb resolution domain data having the lowest AUPRC curves, regardless of predictor type (Supplementary Table 3).

Recursive feature elimination was used to establish the optimal number of distance-type features to include in downstream analyses. It was determined that performances according to AUPRC values were asymptotically similiar for random forest models with sets of predictors greater than 32 (Figure 5). Therefore, we reduced the feature space for subsequent models to the top 32 predictors, respective to resolution and cell line. The sets of top predictors chosen varied across resolution and cell line due to differences in the parameters provided in the TAD calling algorithm.

@@

The low baseline AUPRC values, denoted by the dashed lines in the above PR curves, indicate the severity of

After classifying each genomic bin as either being flanked by a TAD boundary or not, it was found that the subsequent imbalance ratios across resolutions and cell lines were extremely low. Ideally a manageable dataset will have an imbalance ratio hovering near 1. At their heights, the imbalance ratios were 0.179, 0.152, ... for cell lines GM12878, K562, ... respectively (Figure 4).
